---
title: "Homework5"
author: "Zachary Rabow"
date: "March 14, 2019"
output: 
  html_document:
    keep_md: yes
---

```{r}
library('tidyverse')
```

## Mammals Life History
Let's revisit the mammal life history data to practice our `ggplot` skills. Some of the tidy steps will be a repeat from the homework, but it is good practice. The [data](http://esapubs.org/archive/ecol/E084/093/) are from: *S. K. Morgan Ernest. 2003. Life history characteristics of placental non-volant mammals. Ecology 84:3402.*

1. Load the data.
```{r}
life_history <- readr::read_csv("../class_files-master/data/mammal_lifehistories_v2.csv")
```

2. Use your preferred function to have a look. Do you notice any problems?
```{r}
glimpse(life_history)
```

3. There are NA's. How are you going to deal with them?
*There are, I will use a function to replace the "-999's" with actual NA's so they do not skew and alter data calculations*

```{r}
life_history <- 
  life_history %>% 
  na_if("-999")
life_history
```

4. Where are the NA's? This is important to keep in mind as you build plots.

```{r}
life_history %>%
  purrr::map_df(~ sum(is.na(.)))%>% 
  tidyr::gather(variables, num_nas) %>% 
  arrange(desc(num_nas))
```

*The NA's are mostly in wean mass, maximum life, litters per year, weaning count, and AFR observations*

5. Some of the variable names will be problematic. Let's rename them here as a final tidy step.

```{r}
mammals <- 
  life_history %>%  
  dplyr::rename(
          Genus        = Genus ,
          wean_mass    = `wean mass`,
          max_life     = `max. life`,
          litter_size  = `litter size`,
          litters_yr   = `litters/year`
          )
mammals
```

##`ggplot()`
For the questions below, try to use the aesthetics you have learned to make visually appealing and informative plots. Make sure to include labels for the axes and titles.

```{r}
options(scipen=999) #cancels the use of scientific notation for the session
```

6. What is the relationship between newborn body mass and gestation? Make a scatterplot that shows this relationship. 

```{r}
mammals %>% 
  ggplot(aes(x=newborn, y=gestation)) +
  geom_point()+
  scale_x_log10()+
  labs(title = "Newborn Body Mass versus Gestation Time ",
       x = "Newborn Body Mass (grams)",
       y = "Gestation Time (months)")+
  theme(plot.title = element_text(size = rel(1.5)))+
  theme(plot.title = element_text(hjust = 0.5))+
  theme_classic(base_family = "Avenir")
```

7. You should notice that because of the outliers in newborn mass, we need to make some changes. We didn't talk about this in lab, but you can use `scale_x_log10()` as a layer to correct for this issue. This will log transform the y-axis values.

*Ooooops, did that above after the first version looked so gross*

8. Now that you have the basic plot, color the points by taxonomic order.

```{r}
mammals %>% 
  ggplot(aes(x=newborn, y=gestation, color= order)) +
  geom_point()+
  scale_x_log10()+
  labs(title = "Newborn Body Mass versus Gestation Time ",
       x = "Newborn Body Mass (grams)",
       y = "Gestation Time (months)")+
  theme(plot.title = element_text(size = rel(1.5)))+
  theme(plot.title = element_text(hjust = 0.5))+
  theme_classic(base_family = "Avenir")
```

9. Lastly, make the size of the points proportional to body mass.

```{r}
mammals %>% 
  ggplot(aes(x=newborn, y=gestation, color= order, size= mass)) +
  geom_point()+
  scale_x_log10()+
  labs(title = "Newborn Body Mass versus Gestation Time ",
       x = "Newborn Body Mass (grams)",
       y = "Gestation Time (months)")+
  theme(plot.title = element_text(size = rel(1.5)))+
  theme(plot.title = element_text(hjust = 0.5))+
  theme_classic(base_family = "Avenir")+
  scale_size(guide = "none")+
  theme(legend.position = "bottom")
```

10. Make a plot that shows the range of lifespan by order.

```{r}
mammals %>% 
  ggplot(aes(x=order, y=max_life, color=order))+
  geom_boxplot()+
  theme_classic(base_family = "Avenir")+
  labs(title = "Lifespan by  Order",
       x = " Order",
       y = "Lifespan (months)",
       fill="Order")+
  theme(legend.position = "bottom")+
  theme(axis.text.x = element_text(angle=35, hjust=1))+
  theme(plot.title = element_text(hjust = 0.5))
```
