---
title: "Homework3"
author: "Zachary Rabow"
date: "March 14, 2019"
output: 
  html_document:
    keep_md: yes
---

```{r}
fisheries <-  readr::read_csv("../class_files-master/data/FAO_1950to2012_111914.csv")
library(tidyverse)
```

1. Do you see any potential problems with the column names? Does the error message now make more sense?  
```{r}
colnames (fisheries)
```

*Column names are duplicated so R manipulated them to unique names*

2. The `make.names()` command is helpful when there are issues with column names. Notice that although the names are still cumbersome, much of the problemtatic syntax is removed.
```{r}
names(fisheries) = make.names(names(fisheries), unique=T) 
names(fisheries)
```

3. Let's rename the columns. Use `rename()` to adjust the names as follows. Double check to make sure the rename worked correctly. Make sure to replace the old fisheries object with a new one so you can keep the column names.
+ country     = Country..Country.  
+ commname    = Species..ASFIS.species.  
+ sciname     = Species..ASFIS.species._2  
+ spcode      = Species..ASFIS.species._1  
+ spgroup     = Species..ISSCAAP.group.  
+ spgroupname = Species..ISSCAAP.group._1  
+ region      = Fishing.area..FAO.major.fishing.area.  
+ unit        = Measure..Measure.  

renames col.s
```{r}
fisheries_tidy_2 <- 
  rename(fisheries,
         country     = Country..Country.,
         commname    = Species..ASFIS.species., 
         sciname     = Species..ASFIS.species._2, 
         spcode      = Species..ASFIS.species._1,
         spgroup     = Species..ISSCAAP.group.,
         spgroupname = Species..ISSCAAP.group._1,
         region      = Fishing.area..FAO.major.fishing.area.,
         unit        = Measure..Measure.  )
fisheries_tidy_2
```

4. Are these data tidy? Why or why not, and, how do you know?
*The data is not tidy. While each observation has a row, and each value has its own cell, each variabe does not have its own column as the data is currently.*

5. We need to tidy the data using `gather()`. The code below will not run because it is commented (#) out. I have added a bit of code that will prevent you from needing to type in each year from 1950-2012 but you need to complete the remainder `QQQ` and remove the `#`.
```{r}
fisheries_tidy_2 <- 
  fisheries_tidy_2 %>% 
  gather(num_range('X',1950:2012), key='year', value='catch')
```

6. Use `glimpse()` to look at the categories of the variables. Pay particular attention to `year` and `catch`. What do you notice?  
```{r}
glimpse(fisheries_tidy_2)
```
*They are charecter counts and NA values*

7. From question 6 you should see that there are a lot of entries that are missing. In R, these are referred to as NA's but they can be coded in different ways by the scientists in a given study. In order to make the data tidy, we need to deal with them. As a preview to our next lab, run the following code by removing the `#`. It removes the 'X' from the years and changes the `catch` column from a character into a numeric. This forces the blank entries to become NAs. The error "NAs introduced by coercion" indicates their replacement.
```{r}
fisheries_tidy_2 <- 
  fisheries_tidy_2 %>% 
  mutate(
    year= as.numeric(str_replace(year, 'X', '')),
    catch= as.numeric(str_replace(catch, c(' F','...','-'), replacement = '')))
```

8. Are the data tidy? Why?  
```{r}
glimpse(fisheries_tidy_2)
```
*Yes, the data is now tidy as it fixes the prior issue of each variable having a column*

9. You are a fisheries scientist studying cephalopod catch during 2008-2012. Identify the top five consumers (by country) of cephalopods (don't worry about species for now). Restrict the data frame only to our variables of interest.
```{r}
fisheries_tidy_2 %>% 
  select(country, sciname, year, catch) %>% 
  filter (year <=2012 & year >= 2008) %>% 
  filter(sciname== "Cephalopoda") %>% 
  arrange(desc(catch))
```
*Vietnam, India, China, Madagascar, Cambodia*

10. Let's be more specific. Who consumes the most `Common cuttlefish`? Store this as a new object `cuttle`.

```{r}
cuttle <- fisheries_tidy_2
cuttle
```
```{r}
cuttle %>% 
  select(country, commname, year, catch) %>% 
  filter (year <=2012 & year >= 2008) %>% 
  filter(commname== "Common cuttlefish") %>% 
  arrange(desc(catch))
```
*Expanding on Question 9, the countries that consumed Common cuttlefish the most are France, Tunisia, Portugal, and Libya, and Spain*
