---
title: "Homework4"
author: "Zachary Rabow"
date: "March 14, 2019"
output: 
  html_document:
    keep_md: yes
---

```{r}
library('tidyverse')
```

```{r}
life_history <- readr::read_csv("../class_files-master/data/mammal_lifehistories_v2.csv")
```

Rename some of the variables. Notice that I am replacing the old `life_history` data.

```{r}
life_history <- 
  life_history %>% 
  dplyr::rename(
          genus        = Genus,
          wean_mass    = `wean mass`,
          max_life     = `max. life`,
          litter_size  = `litter size`,
          litters_yr   = `litters/year`
          )
```

1. Explore the data using the method that you prefer. Below, I am using a new package called `skimr`. If you want to use this, make sure that it is installed.

```{r}
#install.packages("skimr")
```

```{r}
library("skimr")
```

```{r}
life_history %>% 
  skimr::skim()
```

2. Run the code below. Are there any NA's in the data? Does this seem likely?

```{r}
life_history %>% 
  summarize(number_nas= sum(is.na(life_history)))
```

*There are no NA's but that is unlikely... They are hiding as something else... its my life mission to find them.*

3. Are there any missing data (NA's) represented by different values? How much and where? In which variables do we have the most missing data? Can you think of a reason why so much data are missing in this variable?

*Data NA's are hiden as "-999"'s. Weaning mass data is very unlikely or challenging to collect. like the mass of a weaning bull is probably not the easiest thing to determine*

```{r}
life_history <- 
  life_history %>% 
  na_if("-999")
life_history
```

4. Compared to the `msleep` data, we have better representation among taxa. 
*Produce a summary that shows the number of observations by taxonomic order.*

```{r}
life_history %>% 
  group_by(order) %>% 
  summarize(total=n())
```

5. Mammals have a range of life histories, including lifespan. 

*Produce a summary of lifespan in years by order. Be sure to include the minimum, maximum, mean, standard deviation, and total n.*

```{r}
life_history %>%
  mutate(lifespan=max_life/12) %>% 
  group_by(order) %>%
  summarize(min=min(lifespan, na.rm=TRUE),
            max=max(lifespan, na.rm=TRUE),
            mean=mean(lifespan, na.rm=TRUE),
            sd=sd(lifespan, na.rm=TRUE),
            total=n())
```

6. Let's look closely at gestation and newborns. Summarize the mean gestation, newborn mass, and weaning mass by order. Add a new column that shows mean gestation in years and sort in descending order. Which group has the longest mean gestation? What is the common name for these mammals?

```{r}
life_history%>%
  group_by(order) %>%
 summarize(mean_gestation=mean(gestation, na.rm=TRUE), 
            mean_newborn_mass=mean(newborn, na.rm=TRUE),
            mean_wean_mass=mean(wean_mass, na.rm=TRUE),
            total=n()) %>% 
  arrange(desc(mean_gestation))
```